---
title: "Regex.rmd"
author: "Igor Serafini"
date: "2023-02-06"
output: html_document
---

#Ask if this is right, and Since you picked Species and not genuses?????




#Github link

[My GitHub repository](https://github.com/IgorSerafini/Assignment1_BIOL432)

File name: A5_Serafini_Igor.pdf

###Assignment 5 R-egex

##Part 1

#1

Importing measurements.csv into MData
```{r}
MData<- read.csv("measurements.csv")
```

loading libraries
```{r}
library(dplyr)
```

Inspecting data
```{r}
str(MData)
head(MData)
dim(MData)
```

#2

Adding a new column called Sp that shortens the genus name in the Organism column to only the first letter and period
```{r}
#1st I will create a column that has the same contents as the Species columnn (Genuses) and call it sp. Then I will apply mutate with gsub fufill the preset assignment conditions. 

Sp<- Species_names #Creating object with species names content called sp
  
```

Generating custom gsub function and mutating sp into data frame; to satisfy the following 
  1. Capture First letter
  2. Remainder of the first word = .
```{r}
MData<- MData %>% 
  mutate(Sp = gsub("(\\w).*","\\1.", Sp))

#Lets check our mutated Data

print(MData)
```

Creating new data set with dplyr where the Species_name column (Aka Organism) is substituted by the newly generate Sp column 
```{r}
MData<- MData %>% 
  select(X, Sp, Limb_width, Limb_length, Observer_names)

#Checking the data

print(MData) 
```
Saving the dataset as a new file called MeasShort.csv
```{r}
write.csv(MData, "MeasShort.csv")
```

```{r}

```

