---
title: "A5_Serafini_20232481"
author: "Igor Serafini"
date: "2023-02-08"
output: html_document
---

#####Please Note I had to regenerate my data for week 1, to fit the goal of the assignment#####

Regenerating Data set to include genus names to a file named measurements.csv 
```{r}
set.seed(123)

#Genrate Data for 6 different Genesus 
species_names <- c("Homo sapiens", "Felis catus", "Canis lupus", "Bos taurus", "Equus caballus", "Ursus arctos")

Observer_names <- c("Juan", "James", "Ben", "Ana", "Igor", "Andrei")

# Generate 100 rows of data
measurements <- data.frame(X = 1:100, 
                 Species_names = sample(species_names, 100, replace = TRUE), 
                 Limb_width = round(runif(100, 1.5, 3), 3), 
                 Limb_length = round(runif(100, 1.5, 3), 3), 
                 Observer_names = sample(Observer_names, 100, replace = TRUE))


#Creating .csv measurements data

write.csv(measurements, "measurements.csv")

```

Glossary: 
Homo sapiens (human)
Felis catus (cat)
Canis lupus (gray wolf)
Bos taurus (cattle)
Equus caballus (horse)
Ursus arctos (Bear)


###Assignment 5 R-egex

##Part 1

#Github link

[My GitHub repository](https://github.com/IgorSerafini/Assignment1_BIOL432)

File name: A5_Serafini_Igor.pdf

#1

Importing measurements.csv into MData
```{r}
MData<- read.csv("measurements.csv")
```

loading libraries
```{r}
library(dplyr)
```

Inspecting data
```{r}
str(MData)
head(MData)
dim(MData)
```

#2

Adding a new column called Sp that shortens the genus name in the Organism column to only the first letter and period
```{r}
#1st I will create a column that has the same contents as the Species columnn (Genuses) and call it sp. Then I will apply mutate with gsub fufill the preset assignment conditions. 

Sp<- measurements$Species_names #Creating object with species names content called sp
  
```

Generating custom gsub function and mutating sp into data frame; to satisfy the following 
  1. Capture First letter
  2. Remainder of the first word = .
```{r}
MData<- MData %>% 
  mutate(Sp = sapply(Sp, function(x) gsub("^(\\w)\\w*(\\s)([^ ]+).*","\\1. \\2 \\3", x)))

#Lets check our mutated Data

View(MData)
```

Creating new data set with dplyr where the Species_name column (Aka Organism) is substituted by the newly generate Sp column 
```{r}
MData<- MData %>% 
  select(X, Sp, Limb_width, Limb_length, Observer_names)

#Checking the data

print(MData) 
```
Saving the dataset as a new file called MeasShort.csv
```{r}
write.csv(MData, "MeasShort.csv") #Looks good! Lets commit the code and new data to my Github repository
```




