---
title: "R- Markdown"
author: "Igor Serafini"
date: "2023-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("dataGenerato.R")
source("volumeEstimato.R")
```

GitHub respository: 

[My GitHub repository](https://github.com/IgorSerafini/Assignment1_BIOL432)


#Script 1 (dataGenerato.R):
Generates random values for limb width and length for different species.

#Script 2 (volumeEstimato.R):
Adds a column for volume of a given species leg 

#1
Loading revised measurement.csv data and fixing the extra column (X.1)
```{r}
measurements<- read.csv("measurement.csv")

measurements<- measurements %>% 
  select(-X.1)#removing extra column

head(measurements) #checking data 
```
Loading dplyr  
```{r}
library(dplyr)
```

#2
Sorting Data by species, by observer, and by limb volume: Note, arranged them seperately since the position of each row is relation to all other rows.
```{r}
Sort_Species<- measurements %>% 
  arrange(Species_names) #Sorting by Species 

Sort_Observer<- measurements %>% 
  arrange(Observer_names) #Sorting by Observer  

Sort_Volume<- measurements %>% 
  arrange(Volume) #Sorting by Limb Volume  

#Checking Sort data
print(Sort_Species)
print(Sort_Observer)
print(Sort_Volume)
```
#3
Creating table showing the average volume for each species
```{r}
#First I will create a table with only Species_names and Volume: 

avgVol_Species<- measurements %>% 
  select(-Limb_width, -Limb_length, -Observer_names)

avgVol_Species_table<- avgVol_Species %>% 
  group_by(Species_names) %>% 
  mutate(average_volume= mean(Volume))

#Lets check the data 

print(avgVol_Species_table)

```

#4
Creating table showing the number of observations for each combination of species and observer, by using the group_by() and count() functions 
```{r}
obs_spec<- measurements %>%  
  group_by(Observer_names, Species_names) %>%  
  count()

#Lets check out the data

print(obs_spec)

```

#5
 Creating a single box plot depicting the key features of the distributions of volumes for each species; note, will be loading the ggplot library
```{r}
library(ggplot2)

Plot_VolxSpec<- ggplot(Sort_Species, aes(x= Species_names, y= Volume, fill= Species_names)) + geom_boxplot(outlier.stroke = 0.5) +
  ggtitle("Distributions of Volume for each species") + theme_classic() +
  xlab("Species_names") + ylab("Volume") +
  scale_fill_manual(values = c("#84DCC6", "#D6EDFF", "#F8766DFF", "#8B95C9", "#478978")) 

#Lets check it out 
print(Plot_VolxSpec)
  
```

#6
Creating a multi-panel plot demonstrating a frequency histogram showing distribution of limb volume for each species; Note used coolors to pick the different pallets 
```{r}
Hist_disVol<- ggplot(Sort_Species, aes(x= Volume, fill= Species_names)) + geom_histogram() + facet_wrap(Species_names~.,ncol= 3) + 
  ggtitle("Multi-panel plot demonstrating a frequency histogram of limb volume for each Species") + 
  theme_classic() + scale_fill_manual(values = c("#84DCC6", "#D6EDFF", "#F8766DFF", "#8B95C9", "#478978")) 


#Lets check it out 
print(Hist_disVol) 

```